---
title: "iLINCS API R Notebook"
output:
html_document:
df_print: paged
pdf_document: default
---
<style type="text/css">
  .main-container {
    max-width: 1200px;
    margin-left: auto;
    margin-right: auto;
  }
</style>
### Install iLincsCor 
```{r include = FALSE}
install_from_github <- FALSE
if(install_from_github){
  
devtools::install_github(repo = "Mario-Medvedovic/iLincsCor",force = TRUE)
devtools::install_github(repo = "hrbrmstr/hrbrthemes",force = TRUE)
}
```

### Load Prerequisites  
```{r include = FALSE}
library(ggplot2)
library(hrbrthemes)
library(viridis)
library(knitr)
library(tinytex)
library(httr)
library(jsonlite)
library(htmltools)
library(wCorr)
library(iLincsCor)
options(width = "200")
library(hash)
library(dplyr)
library(weights)
library(rlist)
require(data.table)
library(markdown)
options(max.print=10000)
```

### Define a function to download and retreive ilincs signatures
```{r define function}
get_ilincs_signature <- function(ilincs_signatureId) {
    req <- POST("http://www.ilincs.org/api/ilincsR/downloadSignature", body = list(sigID = paste(ilincs_signatureId), display = FALSE), encode = "json")
    ilincs_sessionId<-unlist(httr::content(req))
    #print(ilincs_sessionId)
    fileUrl=paste("http://www.ilincs.org/tmp/",ilincs_sessionId,".xls",sep="")
    signatureData<-read.table(fileUrl,sep="\t",header=T,stringsAsFactors = F)
    #print(signatureData)
  return(signatureData)
}
```

### Display Signature Libraries
```{r list signature libraries}
apiUrl <- "http://www.ilincs.org/api/SignatureLibraries"
req <- GET(apiUrl)
json <- httr::content(req, as = "text")
ilincs_libraries <- fromJSON(json)
ilincs_libraries[,c("libraryID","libraryName")]
```


#### Select 2 random signature from libraries LIB_5 and check their connectivity against 5 reported connected signature
```{r define parameters}
number_of_random_sample_signatures <- 2
number_of_random_connected_signatures <- 5
# Parameters used for plotting the speed-up
cpu <- c()
time <- c()
run_example <- c()

generate_bing_signature <- FALSE
if(generate_bing_signature){
lincscpDiffExp <- as.matrix(get(load("tests/test_data/lincscpDiffExp_bing.RData")))
lincscpPValExp <- as.matrix(get(load("tests/test_data/lincscpPValues_bing.RData")))

bing_signature <- as.data.frame(cbind(lincscpDiffExp[,1],lincscpPValExp[,1]))
colnames(bing_signature) <- c("Value_LogDiffExp", "Significance_pvalue")
bing_signature['signatureID'] = colnames(lincscpPValExp)[1]
bing_signature['ID_geneid'] = rownames(bing_signature)
bing_signature <- bing_signature[order(as.integer(rownames(bing_signature)),decreasing=FALSE),]
write.table(bing_signature, file='tests/bing_signature.tsv', quote=FALSE, sep='\t', col.names = NA)
}

```

#### Searching for signatures
```{r search signatures}
primes_list <- list("LIB_5")
sig_with_p_value <- c()
sig_without_p_value <- c()

calculated_similarity_with_p_value <- hash()
calculated_similarity_without_p_value <- hash()

ilincs_similarity_with_p_value <- hash()
ilincs_similarity_without_p_value <- hash()

for (p in primes_list) {
  print(p)
 
  ilincs_libId<-p
  apiUrl <- paste("http://www.ilincs.org/api/SignatureMeta/findTermWithSynonyms?library=",ilincs_libId,sep="")
  req <- GET(apiUrl)

  ilincs_result_df<-fromJSON(httr::content(req,type="text"))$data


  print(dim(ilincs_result_df[floor(runif(1000, min=0, max=dim(ilincs_result_df)[1])),c("cellline","compound","concentration","signatureid","is_exemplar")]))
  counter = 0
  seed = 50
  
  for (sigloop in ilincs_result_df[floor(runif(number_of_random_sample_signatures, min=0, max=dim(ilincs_result_df)[1])),c("signatureid")]){
    ilincs_signatureId = sigloop
    print(ilincs_signatureId)
    counter <- counter + 1
    
    print("download signature ======================================")
    req <- POST("http://www.ilincs.org/api/ilincsR/downloadSignature", body = list(sigID = paste(ilincs_signatureId), display = FALSE), encode = "json")
    ilincs_sessionId<-unlist(httr::content(req))
    
    fileUrl=paste("http://www.ilincs.org/tmp/",ilincs_sessionId,".xls",sep="")
    signatureData <-read.table(fileUrl,sep="\t",header=T,stringsAsFactors = F)
  
  if(FALSE){
    apiUrl <- paste("http://www.ilincs.org/api/SignatureMeta/findConcordantSignatures?sigID=",ilincs_signatureId,"&lib=",ilincs_libId,sep="")
    req <- GET(apiUrl)
    
    ilincs_conn_df<-fromJSON(httr::content(req,type="text"))
    result_of_connected_sigs_from_sigID <- ilincs_conn_df[,c("signatureid","similarity","pValue")]
    result_of_connected_sigs_from_sigID<-result_of_connected_sigs_from_sigID[order(result_of_connected_sigs_from_sigID$similarity, decreasing = TRUE),]
    print(head(result_of_connected_sigs_from_sigID))
  }
  
    print("========================================================================")
    print("get connected signatures based on the downloaded signature with p value") 
    print("========================================================================")
    #print(ilincs_sessionId[1])
    #print(toString(ilincs_sessionId[1]))
    
    file_name <- paste(ilincs_libId,"_", toString(ilincs_sessionId[1]), ".tsv", sep="")
    sig_with_p_value <- append(sig_with_p_value,file_name)
    calculated_similarity_with_p_value_list <- list()
    ilincs_similarity_with_p_value_list <- list()


    write.table(signatureData,file=file_name,sep="\t",row.names=F,col.names = T,quote=F)
    #system(paste("head sigFile.tsv"))
    
    
    apiUrl<-"http://www.ilincs.org/api/SignatureMeta/upload"
    sigFile <- file_name
    in_data<-as.data.frame(fread(file_name))
    in_data_orig<-as.data.frame(fread(file_name))
    #print(in_data)
   
    req <- POST(apiUrl, body=list(file=upload_file(sigFile)))
    #req <- POST(apiUrl, body=list(file=signatureData))
    signatureFile <- httr::content(req)$status$fileName[[1]]
    print(head(signatureFile))

    apiUrl <- "http://www.ilincs.org/api/ilincsR/findConcordances"
    req <- (POST(apiUrl, body = list(file=signatureFile, lib=ilincs_libId), encode = "form"))
    output <- data.table::rbindlist(httr::content(req)$concordanceTable, use.names = TRUE, fill = TRUE)
    result_of_upload_complete_sig <- output[,c("signatureid","similarity","pValue")]
    result_of_upload_complete_sig<-result_of_upload_complete_sig[order(result_of_upload_complete_sig$similarity, decreasing = TRUE),]
    print(head(result_of_upload_complete_sig))
    

    print("===========================================================================================================")
    print("checking the accuracy of results via offline benchmarking on the downloaded complete signature with p-value")  
    print("===========================================================================================================")
      for(varIter in 1:number_of_random_connected_signatures)
         {
      
          compared_sig = result_of_upload_complete_sig[varIter,"signatureid"]
        
          restart_data <- in_data
          com_sig <-get_ilincs_signature(toString(compared_sig))
          rownames(com_sig) <- com_sig$ID_geneid
          vardiff <- as.data.frame(com_sig[,c("Value_LogDiffExp")])
          rownames(vardiff) <- com_sig$ID_geneid
          varpval <- as.data.frame(com_sig[,c("Significance_pvalue")])
          rownames(varpval) <- com_sig$ID_geneid
                   
          colnames(vardiff)<- c("compared_sig_dif")
          colnames(varpval)<- c("compared_sig_pval")

          restart_data<-merge(restart_data, vardiff, by.x="ID_geneid", by.y="row.names")
          restart_data<-merge(restart_data, varpval, by.x="ID_geneid", by.y="row.names")
          weight_vec <- c()
          restart_data$"new_weight"<-(-1.0)*(log10(restart_data$"Significance_pvalue")+log10(restart_data$"compared_sig_pval"))
           w_co <- weightedCorr(y=restart_data$"compared_sig_dif", x=restart_data$"Value_LogDiffExp", method="pearson", weights=restart_data$"new_weight")
           
          print(paste0("For signature: ", compared_sig, " calculated weighted correlation is:",w_co, " ilincs query results is: ", result_of_upload_complete_sig[varIter,"similarity"]))
          
          calculated_similarity_with_p_value_list <- append(calculated_similarity_with_p_value_list, w_co)  
          ilincs_similarity_with_p_value_list <- append(ilincs_similarity_with_p_value_list, result_of_upload_complete_sig[varIter,"similarity"])  

      }
    calculated_similarity_with_p_value[[file_name]] <- calculated_similarity_with_p_value_list
    ilincs_similarity_with_p_value[[file_name]] <- ilincs_similarity_with_p_value_list
    
    print("==========================================================================")
    print("get connected signatures based on the downloaded signature without p value") 
    print("==========================================================================")
    
    
    file_name <- paste(ilincs_libId,"_without_pvalue_", toString(ilincs_sessionId[1]), ".tsv", sep="")
    sig_without_p_value <- append(sig_without_p_value, file_name)
      
    calculated_similarity_without_p_value_list  <- list()
    ilincs_similarity_without_p_value_list  <- list()
    
    signatureData_without_pvalue = subset(signatureData, select = -c(Significance_pvalue) )
    write.table(signatureData_without_pvalue,file=file_name,sep="\t",row.names=F,col.names = T,quote=F)
    
    apiUrl<-"http://www.ilincs.org/api/SignatureMeta/upload"
    sigFile <- file_name
    in_data<-as.data.frame(fread(file_name))
    
   
    req <- POST(apiUrl, body=list(file=upload_file(sigFile)))
    
    signatureFile <- httr::content(req)$status$fileName[[1]]
    print(head(signatureFile))

    apiUrl <- "http://www.ilincs.org/api/ilincsR/findConcordances"
    req <- (POST(apiUrl, body = list(file=signatureFile, lib=ilincs_libId), encode = "form"))
    output <- data.table::rbindlist(httr::content(req)$concordanceTable, use.names = TRUE, fill = TRUE)
    result_of_upload_complete_sig <- output[,c("signatureid","similarity","pValue")]
  

    result_of_upload_complete_sig<-result_of_upload_complete_sig[order(result_of_upload_complete_sig$similarity, decreasing = TRUE),]
    print(head(result_of_upload_complete_sig))
    
    
    print("==============================================================================================================")
    print("checking the accuracy of results via offline benchmarking on the downloaded complete signature without p-value")  
    print("==============================================================================================================")
      for(varIter in 1:number_of_random_connected_signatures)
         {
          
          compared_sig = result_of_upload_complete_sig[varIter,"signatureid"]
          
          restart_data <- in_data
          
          
          com_sig <-get_ilincs_signature(toString(compared_sig))
          rownames(com_sig) <- com_sig$ID_geneid
          vardiff <- as.data.frame(com_sig[,c("Value_LogDiffExp")])
          rownames(vardiff) <- com_sig$ID_geneid
          varpval <- as.data.frame(com_sig[,c("Significance_pvalue")])
          rownames(varpval) <- com_sig$ID_geneid
         
          colnames(vardiff)<- c("compared_sig_dif")
          colnames(varpval)<- c("compared_sig_pval")

          restart_data<-merge(restart_data, vardiff, by.x="ID_geneid", by.y="row.names")
          restart_data<-merge(restart_data, varpval, by.x="ID_geneid", by.y="row.names")
          weight_vec <- c()
          restart_data$"new_weight"<-(-1.0)*(log10(restart_data$"compared_sig_pval"))
          w_co <- weightedCorr(y=restart_data$"compared_sig_dif", x=restart_data$"Value_LogDiffExp", method="pearson", weights=restart_data$"new_weight")
           
          print(paste0("For signature: ", compared_sig, " calculated weighted correlation is:",w_co, " ilincs query results is: ", result_of_upload_complete_sig[varIter,"similarity"]))

            calculated_similarity_without_p_value_list <- append(calculated_similarity_without_p_value_list, w_co)  
          ilincs_similarity_without_p_value_list <- append(ilincs_similarity_without_p_value_list, result_of_upload_complete_sig[varIter,"similarity"])  

      }
    calculated_similarity_without_p_value[[file_name]] <- calculated_similarity_without_p_value_list
    ilincs_similarity_without_p_value[[file_name]] <- ilincs_similarity_without_p_value_list
    }
}
for (v in ls(ilincs_similarity_with_p_value)) {
  print(v)
    print(paste(as.list(calculated_similarity_with_p_value[[v]])))
  print(paste(as.list(ilincs_similarity_with_p_value[[v]])))
  print("------------")
}

print("=============================")
for (v in ls(ilincs_similarity_without_p_value)) {
  print(v)
  print(paste(as.list(calculated_similarity_without_p_value[[v]])))
  print(paste(as.list(ilincs_similarity_without_p_value[[v]])))
  print("------------")
}

```

### Running parallel query for comparing the accuracy of parallel query with serial calculation
```{r Running parallel query for LIB_5  with ~1000 genes}

source("tests/create_binary_files_from_lib.R")

create_LIB_fom_file(lib_name="LIB_5",diffExp_data="tests/test_data/lincscpDiffExp.RData"
                                , pPValue_data="tests/test_data/lincscpPValues.RData", seed_sig_csv = sig_with_p_value[1])


# lib <- new(iLincsCor,"data/LIB_test/")
lib <- new(iLincsCor,"data/LIB_5/")
print("Comapring the results of parallel and serial calculations for signatures with p-Value")
first_run <- TRUE
for(item_vec in sig_with_p_value){
  num_process <- 1
  vec <- lib$read_input(item_vec)
  print("=============================")
  print(item_vec)
  print("=============================")
  for(i in seq(1,30)){
  print(paste("number of processors",num_process))
    
    x<-bench::mark(cor_vec <-lib$cor(vec,num_process))
    
    if(first_run){
      time <- append(time,log2(x$min))
      cpu <- append(cpu,log2(num_process))
      run_example <- append(run_example,"LINCS CP library (978 x 143374)")
    }
    
    print("Ordered correlated signatures:")
    print(lib$signature_names[order(cor_vec,decreasing=TRUE)][0:number_of_random_connected_signatures])
    print("Parallel query results:")
    print(cor_vec[order(cor_vec,decreasing=TRUE)][0:number_of_random_connected_signatures])
    print("Serial Weighted Correlation results:")
    print(paste(as.list(calculated_similarity_with_p_value[[item_vec]])))
    print("Serial iLINCS API results:")
    print(paste(as.list(ilincs_similarity_with_p_value[[item_vec]])))
    print("----------------")
    num_process <- i*2
    #print(x)
  }
  first_run <- FALSE
}

print("Comapring the results of parallel and serial calculations for signatures without p-Value")
for(item_vec in sig_without_p_value){
  num_process <- 1
  vec <- lib$read_input(item_vec)
  print("=============================")
  print(item_vec)
  print("=============================")
  iterr = 1
  for(i in seq(1,30)){
  print(paste("number of processors",num_process))
    
    x<-bench::mark(cor_vec <-lib$cor(vec,num_process))
    print("Ordered correlated signatures:")
    print(lib$signature_names[order(cor_vec,decreasing=TRUE)][0:number_of_random_connected_signatures])
    print("Parallel query results:")
    print(cor_vec[order(cor_vec,decreasing=TRUE)][0:number_of_random_connected_signatures])
    print("Serial Weighted Correlation results:")
    print(paste(as.list(calculated_similarity_without_p_value[[item_vec]])))
    print("Serial iLINCS API results:")
    print(paste(as.list(ilincs_similarity_without_p_value[[item_vec]])))
    print("----------------")
    num_process <- i*2
  
    
    
    #num_process <- num_process + 1
    #print(x)
  }
  
}

```



### Running parallel query for a user generated test library
```{r Running parallel query for test}

source("tests/create_test_lib.R")

create_LIB(lib_name="LIB_test", total_sigs=4000000, seed_sig_csv="tests/LINCSCP_1000.tsv")
lib <- new(iLincsCor,"data/LIB_test/")

vec <- lib$read_input("tests/LINCSCP_1000.tsv")

num_process <- 1

benchmark <- TRUE
for(i in seq(1,30)){
  #num_process <- i
  print("========================================")
  print(paste("number of processors", num_process))
  print("========================================")
  if(benchmark){
    x<-bench::mark(cor_vec <-lib$cor(vec,num_process))
    time <- append(time,log2(x$min))
    #print(x)
  }

cor_vec <-lib$cor(vec,num_process)
cpu <- append(cpu,log2(num_process))
run_example <- append(run_example,"test library (978 x 4,000,000)")

num_process <- i*2

}


```



### Running parallel query for speed up of a bigger library
```{r Running parallel query for LIB_5_bing with ~10,000 genes}


source("tests/create_binary_files_from_lib.R")

create_LIB_fom_file(lib_name="LIB_5_bing",diffExp_data="tests/test_data/lincscpDiffExp_bing.RData"
                                , pPValue_data="tests/test_data/lincscpPValues_bing.RData", seed_sig_csv = "tests/bing_signature.tsv")


# lib <- new(iLincsCor,"data/LIB_test/")
lib <- new(iLincsCor,"data/LIB_5_bing/")
vec <- lib$read_input("tests/bing_signature.tsv")

num_process <- 1
for(i in seq(1,30)){
  print(paste("number of processors:",num_process))
  x<-bench::mark(cor_vec <-lib$cor(vec,num_process))
  time <- append(time,log2(x$min))
  cpu <- append(cpu,log2(num_process))
  run_example <- append(run_example,"LINCS CP library (10174 x 143374)")
  num_process <- i*2
  #print(x)
}
```
### Plotting the speed-up
```{r Plotting the speed-up for the 5 tests}

time_series = data.frame(log2_cpu = cpu, log2_time = time, run_example = run_example)

#time_series
print(time_series)
time_series %>%
ggplot( aes(x=cpu, y=time, group=run_example, color=run_example)) +
  geom_line() +
  ggtitle("Time speed-up Curve") +
  xlab("Log2 of Number of CPUs")+
  ylab("Log2 of Computational Time")

```

